A Framework for Equivalence of Cast Calculi and a new Cast Representation

### 2020-10-09 paper.tex
todo in abstract

\begin{document}
%
\title{A Framework for Equivalence of Cast Calculi and a new Cast Representation}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Kuang-Chen Lu\inst{1} \and
Jeremy G. Siek\inst{2} \and
Andre Kuhlenschmidt\inst{2}
}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Brown University, Providence RI 02912, USA \and
Indiana University, Bloomington IN 47405, USA}
% \institute{Princeton University, Princeton NJ 08544, USA \and
% Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
% \email{lncs@springer.com}\\
% \url{http://www.springer.com/gp/computer-science/lncs} \and
% ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
% \email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Gradual typing aims at combining dynamic typing and static typing in one 
language to provide the advantages of both typing disciplines. Implementing 
gradual typing with reasonable overhead, however, is not trivial. Casts might 
accumulate in higher-order values and continuations, thus change the overall 
space complexity. Several space-efficient cast representations have been 
proposed, notably coercions in normal forms and threesomes. But they are
presented in different dynamic semantics, which make it difficult to compare
the key differences between the cast representations. We present a abstrat
data type, namely Cast ADT, that captures interface functions a cast representations
typically includes, and an abstract machine parameterized over Cast ADTs.
The Cast ADT brings another benefit: space-efficient cast representations 
shall be proved not changing the 
overall behavior of language implementations except saving space.
With our generic abstract machine, these proofs can be done by proving properties
of a specific Cast ADT, which is simpler than going through a bisimulation proof 
between a space-efficient dynamic semantics and the definitional dynamic semantics.
We have successfully applied our new approach of proving the correctness of cast 
representations to several known cast representations and believe this framework
is also applicable to other cast representations.
\todo[inline]{JS: cite which other cast representations}
\todo[inline]{KC: will do.}
We also present a new cast representation, hypercoercions, that is more friendly
to mechanized meta-theory and consumes less memory than coercions
in normal forms.

@@

\subsection{Dynamics parameterized over cast representations}
\citet{siek2020mechanized} defines two dynamic semantics for 
cast calculi parameterized over cast representations. The first
semantics, namely CC, is \textit{not} space-efficient and
parameterized over $\mathrm{CastStruct}$. The author instantiates
CC to obtain the blame calculus ($\lambda B$). 
The resulting semantics is analogous to our \LUDMachine.
The other semantics, namely SC, is designed to be space-efficient
and parameterized over $\mathrm{EfficientCastStruct}$. SC is analogous
to our \SMachine{C}. These two parameterized semantics give
users more control over the structure of values. Our \CMachine{B} machine
insists that the cast on the top of a value must be a function cast
or a product cast. In contrast, users of their interfaces define
a predicate on casts called $\mathrm{Inert}$ and any inert cast can
be a part of a value. Their system doesn't include theorems relating
space-inefficient cast calculi and space-efficient ones, which
is one of the major goals of our framework.

@@

\citet{Siek:2010:TWB:1706299.1706342} introduce Lazy UD threesomes and
prove the correctness of Lazy UD threesomes. To use threesomes for
optimization, they define a translation from their cast calculus
(the blame calculus) to a runtime language (threesome calculus).
Then they define a bisimulation relation connecting the blame calculus
and the threesome calculus and prove a bisimulation between them.
In contrast, our \CMachine{B} is an interpreter of the cast calculus and 
compiles casts to cast representations on the fly.

@@

\cite{Siek:2015:BCT:2737924.2737968} proves the correctness
of \LUD\ coercions in normal form. Their approach is similar to
that of \citet{Siek:2010:TWB:1706299.1706342}. They defines
three calculi (i.e. $\lambda B$ (the blame calculus), $\lambda C$, and
$\lambda S$). $\lambda C$ uses coercions but doesn't normalize coercions
so is not space-efficient. $\lambda S$ uses coercions in normal form
and is space-efficient. To establish the correctness of $\lambda S$,
they prove a bisimulation relating $\lambda B$ to $\lambda C$
and a bisimulation relating $\lambda C$ to $\lambda S$.
Inspired by the approaches of \cite{Siek:2010:TWB:1706299.1706342}
and \cite{Siek:2015:BCT:2737924.2737968}, 
we also use the bisimulation technique.

>>>>

\begin{document}
%
\title{A Framework for Equivalence of Cast Calculi and a new Cast Representation}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Kuang-Chen Lu\inst{1} \and
Jeremy G. Siek\inst{2} \and
Andre Kuhlenschmidt\inst{2}
}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Brown University, Providence RI 02912, USA \and
Indiana University, Bloomington IN 47405, USA}
% \institute{Princeton University, Princeton NJ 08544, USA \and
% Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
% \email{lncs@springer.com}\\
% \url{http://www.springer.com/gp/computer-science/lncs} \and
% ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
% \email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Gradual typing aims at combining dynamic typing and static typing in one 
language to provide the advantages of both typing disciplines. Implementing 
gradual typing with reasonable overhead, however, is not trivial. Casts might 
accumulate in higher-order values and continuations, thus change the overall 
space complexity. Several space-efficient cast representations have been 
proposed, notably coercions in normal forms and threesomes. But they are
presented in different dynamic semantics, which make it difficult to compare
the key differences between the cast representations. We present a abstrat
data type, namely Cast ADT, that captures interface functions a cast representations
typically includes, and an abstract machine parameterized over Cast ADTs.
The Cast ADT brings another benefit: space-efficient cast representations 
shall be proved not changing the 
overall behavior of language implementations except saving space.
With our generic abstract machine, these proofs can be done by proving properties
of a specific Cast ADT, which is simpler than going through a bisimulation proof 
between a space-efficient dynamic semantics and the definitional dynamic semantics.
We have successfully applied our new approach of proving the correctness of cast 
representations to several known cast representations
(\cite{Siek:2010:TWB:1706299.1706342,siek2012interpretations,Siek:2015:BCT:2737924.2737968})
and believe this framework
is also applicable to other cast representations (\cite{Garcia:2013:CTB:2500365.2500603,siek2012interpretations}).
\todo[inline]{JS: cite which other cast representations}
\todo[inline]{KC: I am citing \cite{siek2012interpretations} for Lazy D coercions
in normal form, is this the right paper to cite?}
We also present a new cast representation, hypercoercions, that is more friendly
to mechanized meta-theory and consumes less memory than coercions
in normal forms.

@@

\subsection{Dynamics semantics parameterized over cast representations}
\citet{siek2020mechanized} defines two dynamic semantics for 
cast calculi parameterized over cast representations. The first
semantics, namely CC, is \textit{not} space-efficient and
parameterized over $\mathrm{CastStruct}$. The author instantiates
CC to obtain the blame calculus ($\lambda B$) in 
\cite{Siek:2010:TWB:1706299.1706342}. 
The resulting semantics is analogous to our \LUDMachine.
The other semantics, namely SC, is designed to be space-efficient
and parameterized over $\mathrm{EfficientCastStruct}$. SC is analogous
to our \SMachine{C}. CC and SC give their users more control
over the structure of values. Our \CMachine{B} machine
insists that the cast on the top of a value must be a cast between
compound types (specifically, functions and pairs).
In contrast, users of their interfaces define
a predicate on casts called $\mathrm{Inert}$ and any inert cast can
be a part of a value. Their system doesn't include theorems relating
space-inefficient cast calculi and space-efficient ones, which
is one of the major goals of our framework.

@@

\citet{Siek:2010:TWB:1706299.1706342} introduce Lazy UD threesomes and
prove the correctness of this cast representation. To use threesomes for
optimization, they define a translation from their cast calculus
(the blame calculus) to a runtime language (threesome calculus).
Then they prove a bisimulation between the blame calculus
and the threesome calculus.
In contrast, our \CMachine{B} is an interpreter of the cast calculus and 
compiles casts to cast representations on the fly.

@@

\cite{Siek:2015:BCT:2737924.2737968} proves the correctness
of \LUD\ coercions in normal form. Their approach is similar to
that of \citet{Siek:2010:TWB:1706299.1706342}. They defines
three calculi (i.e. $\lambda B$ and $\lambda C$ and $\lambda S$). 
$\lambda B$ is the blame calculus. 
$\lambda C$ uses coercions but doesn't normalize coercions
so is not space-efficient. $\lambda S$ uses coercions in normal form
and is space-efficient. To establish the correctness of
their cast representation,
they prove a bisimulation relating $\lambda B$ to $\lambda C$
and a bisimulation relating $\lambda C$ to $\lambda S$.
Inspired by the approaches of \cite{Siek:2010:TWB:1706299.1706342}
and \cite{Siek:2015:BCT:2737924.2737968}, 
we also use the bisimulation technique.
